---
title: "Untitled"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# tidyroc

This is a repository for an R package `tidyroc`. `tidyroc` is currently under development, and I plan to release it this summer (summer 2019). 

## Usage

`tidyroc` has 3 primary functions: 

* to generate values to plot a receiver operating characteristic (ROC) curve  
* to generate values to plot a precision-recall curve  
* to calculate area under an ROC and precision-recall curves (AUC)

`tidyroc` is designed to be integrated with the Tidyverse. It is intended to work with `broom`, `dplyr`, and `ggplot2`.

## Examples

#### Plot ROC curves for 2 different models
```{r, message = FALSE}
# load packages required to run this code chunk
library(ggplot2)
library(dplyr)
library(broom)
library(cowplot)

# load tidyroc
devtools::load_all(".")

# make a reduced iris data set that only contains virginica and versicolor species
iris.small <- filter(iris, Species %in% c("virginica", "versicolor"))

# fit a logistic regression model to the data to assess the model's performance
d1 <- glm(Species ~ Petal.Width + Petal.Length + Sepal.Width,
  data = iris.small,
  family = binomial
) %>%
  augment() %>%
  mutate(model = "model1") # name the model

# fit a different logistic regression model to the data to assess this model's performance
d2 <- glm(Species ~ Petal.Width,
  data = iris.small,
  family = binomial
) %>%
  augment() %>%
  mutate(model = "model2") # name the model

# combine the two datasets to make an ROC curve for each model
d <- bind_rows(d1, d2)

# plot ROC curves
d %>%
  group_by(model) %>% # group to get individual ROC curve for each model
  make_roc(predictor = .fitted, known_class = Species) %>% # get values to plot an ROC curve
  ggplot(aes(x = fpr, y = tpr, color = model)) +
  geom_line(size = 1.1) +
  scale_color_manual(values = c("#C04A56", "#3D8CF1")) +
  theme_cowplot()
```


#### Plot precision-recall curves for 2 different models
```{r, message=FALSE}
# plot precision-recall curves using the data-frame we generated in the previous example
d %>%
  group_by(model) %>% # group to get individual precision-recall curve for each model
  make_pr(predictor = .fitted, known_class = Species) %>% # get values to plot a precision-recall curve
  ggplot(aes(x = recall, y = precision, color = model)) +
  geom_line(size = 1.1) +
  coord_cartesian(ylim = c(0,1), xlim = c(0,1)) +
  scale_color_manual(values = c("#C04A56", "#3D8CF1")) +
  theme_cowplot()
```

