---
title: "tidyroc"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "figures/"
)
```
# tidyroc

This is a repository for an R package `tidyroc`. `tidyroc` is currently under development, and I plan to release it this summer (summer 2019). 

## Usage

`tidyroc` has 3 primary functions: 

* `make_roc()` calculates true positive and false positive rates to plot a receiver operating characteristic (ROC) curve
* `make_pr()` calculates presicion and recall (true positive rate) to plot a precision-recall curve  
* `calc_auc()` calculates the area under an ROC (AUC)

`tidyroc` is designed to be integrated with the Tidyverse. It is intended to work with `broom`, `dplyr`, and `ggplot2`. Here is a simple use case.
```{r, eval = FALSE}
glm(outcome ~ predictor1 + predictor2 + predictor3, # fit a model using 3 predictors
  data = df,
  family = binomial
) %>%
  augment() %>% # use broom to add glm output to the original data frame
  make_roc(predictor = .fitted, known_class = Species) %>% # get values to plot an ROC curve
  ggplot(aes(x = fpr, y = tpr, color = model)) + # plot false positive rate against true positive rate
  geom_line(size = 1.1)
```


## Examples

Before I make any plots, I fit two logitistic regression models. I will use the `iris` dataset for this purpose. First, I reduce `iris` to only contain data for two species of irises. This ensures that the outcome is binary. 

I load the libraries needed to run the examples below.
```{r, message = FALSE}
# load tidyverse packages
library(ggplot2)
library(dplyr)
library(broom)

# load cowplot to change plot theme
library(cowplot)

# load tidyroc
devtools::load_all(".")
```

I reduce the `iris` dataset to contain information only about virginica and versicolor. I fit two logistic regression models that attempt to predict the two species of irises. The first model tries uses petal width, petal length, and sepal width as predictors. The second model uses only sepal width as a predictor.

```{r, message = FALSE}
# reduce iris to two species to fit a model that is a binary classifier
iris.small <- filter(iris, Species %in% c("virginica", "versicolor"))

# fit a logistic regression model to the data to assess the model's performance
glm.out1 <- glm(Species ~ Petal.Width + Petal.Length + Sepal.Width,
  data = iris.small,
  family = binomial
) %>%
  augment() %>%
  mutate(model = "model1") # name the model

# fit a different logistic regression model to the data to assess this model's performance
glm.out2 <- glm(Species ~ Sepal.Width,
  data = iris.small,
  family = binomial
) %>%
  augment() %>%
  mutate(model = "model2") # name the model

# combine the two datasets to make an ROC curve for each model
glm.out <- bind_rows(glm.out1, glm.out2)
```
Now that I have fitted values, I can make a plot with two ROC curves and a plot with two precision-recall curves. I can also calculate the area under each of the ROC curves.

#### Plot ROC curves for 2 different models
```{r, message = FALSE}
# plot ROC curves
glm.out %>%
  group_by(model) %>% # group to get individual ROC curve for each model
  make_roc(predictor = .fitted, known_class = Species) %>% # get values to plot an ROC curve
  ggplot(aes(x = fpr, y = tpr, color = model)) +
  geom_line(size = 1.1) +
  geom_abline(slope = 1, intercept = 0, size = 0.4) + 
  scale_color_manual(values = c("#001889", "#AB1488")) +
  theme_cowplot()
```


#### Plot precision-recall curves for 2 different models
```{r, message=FALSE}
# plot precision-recall curves using the data-frame we generated in the previous example
glm.out %>%
  group_by(model) %>% # group to get individual precision-recall curve for each model
  make_pr(predictor = .fitted, known_class = Species) %>% # get values to plot a precision-recall curve
  ggplot(aes(x = recall, y = precision, color = model)) +
  geom_line(size = 1.1) +
  coord_cartesian(ylim = c(0,1), xlim = c(0,1)) +
  scale_color_manual(values = c("#001889", "#AB1488")) +
  theme_cowplot()
```

#### Calculate AUC of two ROC curves
```{r}
glm.out %>%
  group_by(model) %>% # group to get individual precision-recall curve for each model
  make_roc(predictor = .fitted, known_class = Species) %>% 
  summarise(auc = calc_auc(x = fpr, y = tpr))
```

